---
title: 'Onaq Summer 2020: Precipitation, VSWC, Latent Heat'
author: "Paige Austin"
date: "3/8/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
Goals for 5/2020 through 7/2020:

1. Load: precip, LE, Stemp, VSWC, and precip d18O
2. Align: start date/time
3. Aggregate: 12-hour time blocks

```{r setup, echo = FALSE, results='hide', message=FALSE}
knitr::opts_chunk$set(error = TRUE)
#setwd("~/Documents/Projects/inversion/data")
library(neonUtilities)
library(ggplot2)
library(dplyr)
library(imputeTS) #gg_na_dist, statsNA
library(timetk) #filter_by_time
library(knitr)
library(ggpubr) #ggarrange
#BiocManager::install('rhdf5') #ec data

options(stringsAsFactors=F)
```


Load data:

Precipitation: DP1.00006.001 (units: mm)

Eddy Covariance Bundle: DP4.00200.001 (units: Wm-2)

Volumetric Soil Water Content: DP1.00094.001 (unitless)

```{r pcp_loading,results = "hide" }
#loading 5 minute data from April - August 2020, need May - July 2020
pcp <- loadByProduct(dpID="DP1.00006.001",
                           site= "ONAQ",
                           timeIndex = "5",
                           startdate="2020-04",
                           enddate="2020-08",
                           check.size = F)
```

Precip. data prep
```{r pcp_prep}
#5 min data df for 4/1/20 to 8/31/20
pcp_5min <- pcp$PRIPRE_5min

#filtering by time for May to July 2020
pcp_0507 <-pcp_5min %>%
    filter_by_time(.date_var = endDateTime,
                   .start_date = "2020-05-01 00:00:00",
                   .end_date = "2020-07-31 23:55:00") 

#only quality filtered data
pcp_0507_qc <- pcp_0507 %>%
      filter(priPrecipFinalQF == "0")

```

EC bundle
```{r ec_loading, eval = FALSE}
# #loading 30 minute ONAQ eddy covariance data from May to July 2020
#   zipsByProduct(dpID = "DP4.00200.001",
#   site = "ONAQ",
#   startdate = "2020-05",
#   enddate = "2020-07",
#   package = "basic",
#   check.size = F,
#   savepath = "./onaq_0507_2020") #change file path for storing data

```
Reading downloaded data: processed fluxes, level 4 data product
```{r ec_read, results = "hide"}
#pullling out 30 min calculated fluxes
#use filepath specified above
flux <- stackEddy(filepath = "./onaq_0507_2020/filesToStack00200",
                 level="dp04") 

obj_desc <- flux$objDesc

variables <- flux$variables

```

EC data prep
```{r ec_prep}
#pull out dataframe of all fluxes
onaq <- flux$ONAQ

#making dataframe by unflagged LE data
onaq.qc <- onaq %>%
        filter(qfqm.fluxH2o.nsae.qfFinl == "0")
```

VSWC loading
```{r vswc_loading,results = "hide", message= FALSE}
#loading 30 minute ONAQ somo data from April - August 2020, need May - July
swc_0408 <- loadByProduct(dpID="DP1.00094.001",
                           site= "ONAQ",
                           timeIndex = "30",
                           startdate="2020-04",
                           enddate="2020-08",
                           check.size = F)

#30 min data df for 4/1/20 to 8/31/20
swc_0408 <- swc_0408$SWS_30_minute

#filtering for May - July 2020 and by unflagged data
swc_0507_qc <- swc_0408 %>%
    filter_by_time(.date_var = endDateTime,
                   .start_date = "2020-04-30 23:30:00",
                   .end_date = "2020-08-01 00:00:00") %>%
    filter(VSWCFinalQF == "0")

#filtering by h-pos 5, v-pos 501; h-pos 2, v-pos 502, 503
swc_0507_go <-  swc_0507_qc %>% 
        filter(horizontalPosition == "005" &
               verticalPosition == "501" |
        horizontalPosition == "002" & verticalPosition == "502" |
        horizontalPosition == "002" & verticalPosition == "503" |
        horizontalPosition == "002" & verticalPosition == "504" |
        horizontalPosition == "001" & verticalPosition == "505")


swc_0507_go %>%
        group_by(horizontalPosition, verticalPosition) %>%
        ggplot(aes(x= endDateTime,y= VSWCMean, 
                   color = verticalPosition)) +
                     geom_line() + labs(y='Mean VSWC', x = 'Time') +
        facet_wrap(vars(horizontalPosition)) + theme_light() +
        ggtitle("Quality Filtered 30 min. ONAQ Mean VSWC")

```

soil temp loading
```{r stemp_loading,results = "hide", message= FALSE}
#loading 30 minute ONAQ stemp data from April - August 2020, need May - July
t_0408 <- loadByProduct(dpID="DP1.00041.001",
                           site= "ONAQ",
                           timeIndex = "30",
                           startdate="2020-04",
                           enddate="2020-08",
                           check.size = F)

#30 min data df for 4/1/20 to 8/31/20
t_0408 <- t_0408$ST_30_minute

#filtering for May - July 2020 and by unflagged data
t_0507_qc <- t_0408 %>%
    filter_by_time(.date_var = endDateTime,
                   .start_date = "2020-04-30 23:30:00",
                   .end_date = "2020-08-01 00:00:00") %>%
    filter(finalQF == "0")

t_0507_qc %>%
        group_by(horizontalPosition, verticalPosition) %>%
        ggplot(aes(x= endDateTime,y= soilTempMean, 
                   color = verticalPosition)) +
                     geom_line() + labs(y='Mean Stemp', x = 'Time') +
        facet_wrap(vars(horizontalPosition)) + theme_light() +
        ggtitle("Quality Filtered 30 min. ONAQ Mean Stemp")

```

relative humidity loading
```{r rh_loading,results = "hide", message= FALSE}
#loading 30 minute ONAQ stemp data from April - August 2020, need May - July
rh_0408 <- loadByProduct(dpID="DP1.00098.001",
                           site= "ONAQ",
                           timeIndex = "30",
                           startdate="2020-04",
                           enddate="2020-08",
                           check.size = F)

#30 min data df for 4/1/20 to 8/31/20
rh_0408 <- rh_0408$RH_30min

#filtering for May - July 2020 and by unflagged data
rh_0507_qc <- rh_0408 %>%
    filter_by_time(.date_var = endDateTime,
                   .start_date = "2020-04-30 23:30:00",
                   .end_date = "2020-08-01 00:00:00") %>%
    filter(RHFinalQF == "0", verticalPosition == "000")

rh_0507_qc %>%
        ggplot(aes(x = endDateTime, y = RHMean, color = verticalPosition)) +
          geom_line() + labs(y='RH Mean', x = 'Time') + theme_light() +
          ggtitle("Quality Filtered 30 min. ONAQ Mean Surface RH")

```

```{r}
#DICEE downscaled precipitation files downloaded from:
#https://www.hydroshare.org/resource/e74edc35d45441579d51286ea01b519f/
dop.m = read.csv("bigdata/daily_p_d18O_mean.csv")
dop.sd = read.csv("bigdata/daily_p_d18O_std.csv")
#Keep just ONAQ
dop = data.frame(date = as.POSIXct(dop.m$Date), d18O = dop.m$ONAQ, d18OSD = dop.sd$ONAQ)
#Remove NAs
dop = dop[dop$d18O != -9999,]
#Parse daily values to 12-hour time series using nearest neighbor
dop = dop[dop$date > "2020-04-30" & dop$date < "2020-08-01",]
p = read.csv("data/pcp_onaq_12hour_exdset3_05072020.csv")
ts = as.POSIXct(p$t)
ts.do.m = ts.do.sd = numeric(length(ts))
for(i in seq_along(ts)){
  toff = abs(ts[i] - dop$date)
  tind = match(min(toff), toff)
  ts.do.m[i] = dop$d18O[tind]
  ts.do.sd[i] = dop$d18OSD[tind]
}
dop = data.frame("dateTime" = ts, "d18O.m" = ts.do.m, 
                 "d18O.sd" = ts.do.sd)
```

```{r}
#Calibrated ONAQ water vapor d18O from Rich
wv = read.csv("bigdata/ONAQ_wiso_2020.csv")
wv$timeEnd = as.POSIXct(wv$timeEnd, format = "%Y-%m-%d %H:%M:%S")
wv$timeBgn = as.POSIXct(wv$timeBgn, format = "%Y-%m-%d %H:%M:%S")

#filtering for May - July 2020 and by unflagged data
wv_0507 <- wv %>%
    filter_by_time(.date_var = timeEnd,
                   .start_date = "2020-04-30 23:59:59",
                   .end_date = "2020-08-01 00:00:00")

```


Aligning timeseries: filtering by start time (GMT)

GMT is 7 hours ahead

Mountain standard time 8:00am = 15:00 GMT same day

Mountain standard time 8:00pm = 03:00 GMT next day
```{r align}

#filtering start times beginning 03:00 GMT May 05 (8:00pm MST)
#filtering for May 05 - July 2020 
ec_aligned <- onaq.qc %>%
    filter_by_time(.date_var = timeBgn,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

swc_aligned <- swc_0507_go %>%
    filter_by_time(.date_var = startDateTime,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

t_aligned <- t_0507_qc %>%
    filter_by_time(.date_var = startDateTime,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

pcp_aligned <- pcp_0507_qc %>%
    filter_by_time(.date_var = startDateTime,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

wv_aligned = wv_0507 %>%
  filter_by_time(.date_var = timeBgn,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

rh_aligned = rh_0507_qc %>%
  filter_by_time(.date_var = startDateTime,
                   .start_date = "2020-05-05 03:00:00",
                   .end_date = "2020-08-01 00:00:00")

```

Removing constant drizzle from precip. data
```{r pcp_dp}
#viz
p1 <- ggplot(pcp_aligned, aes(x= endDateTime, y= priPrecipBulk)) +
                geom_line() +
                geom_hline(yintercept = 0.1, color = "red") +
                labs(y = "Precipitation (mm)", x = "Time") +
                ggtitle("ONAQ Bulk Precip. May - July, 2020",
  subtitle = " Quality filtered 5 min data with suggested drizzle <= 0.1")
p1

#Checking for constant drizzle
density(pcp_aligned$priPrecipBulk)
plot(density(pcp_aligned$priPrecipBulk), 
     xlim = c(-0.01, 0.5), ylim = c(0, 1.5),
     main = "Quality Filtered Primary Precip.")
abline(v = 0.2, col = "blue")

#try 0.2 for cutoff
#removing drizzle, setting everything less than and = 0.2 to zero
pcp_aligned$bulk_nodrizz <- ifelse(pcp_aligned$priPrecipBulk > 0.2, pcp_aligned$priPrecipBulk, 0)

#viz: time and quality filtered, drizzle removed
p3 <- ggplot(pcp_aligned, aes(x= endDateTime, y= bulk_nodrizz)) +
              geom_line() +
              geom_hline(yintercept = 0.2, color = "red") +
              labs(y = "Precipitation (mm)", x = "Time") +
            ggtitle("Quality Filtered ONAQ Bulk Precip. May - July, 2020",
    subtitle = "5 min data with suggested drizzle <= 0.2 mm set to zero")

p3

```

Time aggregation to 12-hour blocks
```{r time_agg}
#aggregating by time
#add a col for the time aggregation, set to datetime 
#summarize by sum pcp, count records, count missing, calc. percent complete  
pcp_12h <- pcp_aligned %>%
            mutate(t = as.POSIXct(cut(endDateTime, breaks = "12 hour")),
                "%Y-%m-%d %H:%M:%S", tz = "GMT")  %>%
            group_by(t) %>% #group by time
            summarise(bulk_precip = sum(bulk_nodrizz, na.rm = TRUE),
            n = n(),
            na = sum(is.na(bulk_nodrizz)),
            percent_complete = 1 - sum(is.na(bulk_nodrizz))/n())

#viz: 12 hr aggregated with 0.2 drizzle removed
p4 <- ggplot() +
          geom_line(data = pcp_12h, aes(x= t, y= bulk_precip)) +
          labs(y = "Precipitation (mm)", x = "Time") +
          ggtitle("12 Hr. ONAQ Precip. May - July, 2020",
          subtitle = "0.2 mm drizzle removed")
p4

#time aggregation to 12 hour blocks, even with missing data
ec_12h <- ec_aligned %>%
      mutate(t = as.POSIXct((cut(timeEnd, breaks = "12 hour")),
                            "%Y-%m-%d %H:%M:%S", tz = "GMT")) %>%
      group_by(t) %>% #group by time interval
      summarise(mean_nsae_h2o = mean(data.fluxH2o.nsae.flux, na.rm = T), 
      n = n(),
      na = sum(is.na(data.fluxH2o.nsae.flux)),
      percent_complete = 1 - sum(is.na(data.fluxH2o.nsae.flux))/n())

#plotting filtered 12hr aggregated H2O NSAE 
p5 <- ec_12h %>%
        ggplot(aes(x= t,y = mean_nsae_h2o)) +
                geom_point() +
                geom_line() + labs(x = "Time", y ='NSAE H2O Flux (Wm-2)') +
                theme_light() +
        ggtitle(label = "Quality Filtered NSAE H2O Flux",
        subtitle = "Time Aggregated to 12 Hour Mean")
p5

#keeping h-pos, v-pos and aggregating by time
#add a col for the time aggregation, set to datetime 
#group by h-pos, v-pos and time
#summarize by mean vswc 
##first partition uncertainty into calibration (fixed) and measurement error (variable)
swc_aligned$VSWCCalUncert = rep(min(swc_aligned$VSWCExpUncert) / 2)
swc_12h <- swc_aligned %>%
      mutate(t = as.POSIXct((cut(endDateTime, breaks = "12 hour")),
                            "%Y-%m-%d %H:%M:%S", tz = "GMT")) %>%
      group_by(horizontalPosition, verticalPosition, t) %>% 
      summarise(mean_vswc = mean(VSWCMean, na.rm = T), 
                cal_uncert = mean(VSWCCalUncert, na.rm = TRUE),
                meas_uncert = sd(VSWCMean, na.rm = TRUE) / sqrt(n()),
      n = n(),
      na = sum(is.na(VSWCMean)),
      percent_complete = 1 - sum(is.na(VSWCMean))/n())
mum = max(swc_12h$meas_uncert, na.rm = TRUE)
swc_12h$meas_uncert[swc_12h$meas_uncert == 0] = mum
swc_12h$meas_uncert[is.na(swc_12h$meas_uncert)] = mum

p6 <- swc_12h %>%
        group_by(horizontalPosition, verticalPosition) %>%
        ggplot(aes(x= t,y= mean_vswc, 
                   color = verticalPosition)) +
                   geom_line() + labs(y='Mean VSWC') +
        facet_wrap(vars(horizontalPosition)) + theme_light() +
        ggtitle("Quality Filtered 12 Hr. ONAQ Mean VSWC") 
p6

#keeping h-pos, v-pos and aggregating by time
#add a col for the time aggregation, set to datetime 
#group by time, keep only v-pos 501
#summarize by mean stemp
t_aligned = t_aligned[t_aligned$verticalPosition == "501",]
t_12h <- t_aligned %>%
      mutate(t = as.POSIXct((cut(endDateTime, breaks = "12 hour")),
                            "%Y-%m-%d %H:%M:%S", tz = "GMT")) %>%
      group_by(t) %>%
      summarise(mean_temp = mean(soilTempMean, na.rm = T), 
      n = n(),
      na = sum(is.na(soilTempMean)),
      percent_complete = 1 - sum(is.na(soilTempMean))/n())

p7 <- t_12h %>%
        ggplot(aes(x= t,y= mean_temp)) +
                   geom_line() + labs(y='Mean Stemp') + theme_light() +
        ggtitle("Quality Filtered 12 Hr. ONAQ Mean Stemp") 
p7

#keeping v-pos and aggregating by time
#add a col for the time aggregation, set to datetime 
#group by h-pos, v-pos and time
#summarize by mean vswc 
wv_12h <- wv_aligned %>%
      mutate(t = as.POSIXct((cut(timeEnd, breaks = "12 hour")),
                            "%Y-%m-%d %H:%M:%S", tz = "GMT")) %>%
      group_by(level, t) %>% 
      summarise(d18O_m = mean(d18O, na.rm = TRUE),
                d18O_sd = sd(d18O, na.rm = TRUE) / sqrt(n()),
                d2H_m = mean(d2H, na.rm = TRUE),
                d2H_sd = sd(d2H, na.rm = TRUE) / sqrt(n()),
                sh_m = mean(spec_hum, na.rm = TRUE),
                sh_sd = sd(spec_hum, na.rm = TRUE) / sqrt(n()),
                n = n())

p8 <- wv_12h %>%
        group_by(level) %>%
        ggplot(aes(x= t,y= d18O_m, 
                   color = level)) +
                   geom_line() + labs(y='Mean d18O') +
        facet_wrap(vars(level)) + theme_light() +
        ggtitle("Quality Filtered 12 Hr. ONAQ Mean VSWC") 
p8

#aggregating by time
#add a col for the time aggregation, set to datetime 
#group by time summarize by mean RH
rh_12h <- rh_aligned %>%
      mutate(t = as.POSIXct((cut(endDateTime, breaks = "12 hour")),
                            "%Y-%m-%d %H:%M:%S", tz = "GMT")) %>%
      group_by(t) %>%
      summarise(mean_rh = mean(RHMean, na.rm = TRUE), 
      n = n(),
      na = sum(is.na(RHMean)),
      percent_complete = 1 - sum(is.na(RHMean))/n())

p9 <- rh_12h %>%
        ggplot(aes(x = t,y = mean_rh)) +
                   geom_line() + labs(y='Mean RH') + theme_light() +
        ggtitle("Quality Filtered 12 Hr. ONAQ Mean Surface RH") 
p9

```

Data Quality for LE, about 35% missing data
```{r dq}
p7 <- ec_12h %>%
        ggplot(aes(x= t,y = mean_nsae_h2o, color = percent_complete)) +
                geom_point() +
                geom_line() + labs(x = "Time", y ='NSAE H2O Flux (Wm-2)') +
                theme_light() +
        ggtitle(label = "Quality Filtered NSAE H2O Flux",
        subtitle = "Time Aggregated to 12 Hour Mean") 
       
p7

```
Writing file for Summer 2020 data

VSWC data:

horizontal position 1, vertical position 505
horizontal position 2, vertical positions 502-504
horizontal position 5, vertical positions 501


```{r, eval = FALSE, echo = FALSE}
#writing file for 12 hour LE data for 05/20 to 07/20
#write.csv(ec_12h, file = "le_nsae_onaq_12hour_exdset2_05072020.csv")

#writing file for 12 hour VSWC data for 05/20 to 07/20
write.csv(swc_12h, file = "data/swc_onaq_12hour_exdset3_05072020.csv")

#writing file for 12 hour Stemp data for 05/20 to 07/20
write.csv(t_12h, file = "data/st_onaq_12hour_05072020.csv")

#writing file for 12 hour precip d18O data for 05/20 to 07/20
write.csv(dop, file = "data/d18Op_onaq_12hour_05072020.csv")

#writing file for 12 hour wv data for 05/20 to 07/20
write.csv(wv_12h, file = "data/wvIso_onaq_12hour_05072020.csv")

#writing file for 12 hour wv data for 05/20 to 07/20
write.csv(rh_12h, file = "data/rh_onaq_12hour_05072020.csv")

#writing file for 12 hour Precip. data for 05/20 to 07/20
#write.csv(pcp_12h, file = "pcp_onaq_12hour_exdset2_05072020.csv")
```
